RadLib: a radiative property model library for combustion CFD

### 2021-07-20 submissions/CPC_response_to_reviewers.md

2. Page 10, missing information on [INSERT RUNTIME COMPARISONS HERE].

	**This was a placeholder statement that should not have been included in the document. It has been removed.**

3. In Figure 2, the RCSLM seems to consistently perform better than WSGG, except in Example S2. In addition, WSGG performs worse than PM in Example S4. Can the authors explain why?

    ** The WSGG model coefficients presented were those from Bordbar's paper. These have since been updated and included in the revised code and we have likewise updated the simulation results in the paper. For the new results, the WSGG model no longer performs poorly for example S4. For example S2, the RCSLW model converges to the LBL results as the number of gases increases, but the WSGG model outperforms the RCSLW model for 4 gray gases. It is not clear why the WSGG model outperforms the RCSLW model for this case. We have added some relevant discussion to the paper. We do note that the WSGG model is empirical and contains very little representative physics, which makes it difficult to explain the behavior of the model.**

### 2021-07-20 manuscript/manuscript.tex
DOL: edits to the results section; updating figures with revised wsgg coefficients; modified response to reviewers

Weighted sum of gray gases (WSGG) models generate radiation absorption coefficients for use with the spectrally integrated RTE (Eq.~\ref{e:RTE_wsgg}).
RadLib implements the WSGG model presented by Bordbar et al. \citep{Bordbar_2014,Bordbar_2020}, which uses correlations based on the HITEMP 2010 database \cite{Rothman_2010}. It models mixtures of gaseous \ce{CO2} and \ce{H2O}, the two most significant radiatively participating species in most combustion environments, using four gray gases and one clear gas. Absorption coefficients are calculated by 
%
\begin{linenomath}
\begin{equation}
    \kappa_j=\sum_{k=0}^{4}d_{j,k}M_r^k,
\end{equation}
\end{linenomath}
%
where $\kappa_j$ is the absorption coefficient for gray gas $j$, $d_{j,k}$ is a correlated model coefficient, and $M_r$ is the molar ratio $y_{\mathrm{H_2O}}/y_{\mathrm{CO_2}}$. The weight factors are calculated by 
%
\begin{linenomath}
\begin{equation}
    a_j=\sum_{k=0}^{4}b_{j,k}T_r^k,
\end{equation}
\end{linenomath}
%
where $a_j$ is the weighting factor for gas $j$ and $T_r$ is a scaled temperature equal to $T/T_{ref}$ with $T_{ref}=1200$K. The valid temperature range of the model is 300-2400 K; temperatures outside this range use the value at the respective bound. The value of $b_{i,j}$ is calculated by 
%
\begin{linenomath}
\begin{equation}
 b_{j,k}=\sum_{i=0}^{4}C_{j,k,i}M_r^i,
\end{equation}
\end{linenomath}
%
where $C_{j,k,i}$ is another correlated model coefficient. The model coefficients $d_{j,k}$ and $C_{j,k,i}$ are provided by Bordbar et~al. \citep{Bordbar_2014,Bordbar_2020}.

@@

Table~\ref{t:examples} summarizes the example cases. Each case is labeled according to its source---Solvojov et al. \cite{Solovjov_2017} (S), Bordbar et al. \cite{Bordbar_2020} (B), or Solovjov et~al. \cite{Solovjov_2001} (Sb)---and the number of each example corresponds to the example number in its respective reference. Each example case consists of one or more gas layers ("slabs") bounded by black walls. 
Example S1 consists of two adjacent gas layers with identical species compositions: one hot layer of constant thickness and one cold layer of varying thickness. 
In Example S2, the two gas layers are isothermal but differ in the value of the \ce{CO2} mole fraction, where the layer thickness of the "thin" slab (with the smaller \ce{CO2} mole fraction) varies.
Example S3 applies parabolic temperature and \ce{H2O} mole fraction profiles across a single gas layer. 
Example S4 divides the gas into three equally-spaced layers and applies a triangular temperature profile to the middle layer and a constant temperature to the two outer layers.
Example S5 uses a half-sinusoid temperature profile that decreases from 1500 to 500 K applied to a single gas layer of constant species composition.
Example B3 has symmetric temperature and H$_2$O mole fraction profiles with central peaks of 1800 K and 1, respectively (with $y_{CO2}=1-y_{H2O}$). 
Example Sb1 consists of a single gas with a uniform temperature and composition profile and soot radiation. 
%
\begin{table}
    \caption{Summary of example cases. All cases use $P=1$ atm and black walls.}
    \label{t:examples}
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c l l c c c c}
        \hline
        Example & T(K)                          & $y_{H2O}$  $y_{CO2}$ (mole frac.)               & L (m)   & $T_{walls}$ (K) & Ref.\\
        \hline
        S1 & $T(x<0.5)=2000$; $T(x>0.5)=300$    & $y_{CO2}=0.1$, $y_{H2O}=0.2$                    & 0.5-2.5 & cold, cold	& \cite{Solovjov_2017}     \\
        S2 & T=1000                             & $y_{CO2}(x<0.5)=0.4$, $y_{CO2}(x>0.5)=0.1$      & 0.5-2.5 & cold, cold & \cite{Solovjov_2017}    \\
           &                                    & $y_{H2O}=0.0$                                   &         &                \\
        S3 & $T(x) = 4000x(L-x)/L^2 + 800$      & $y_{H2O}(x) = 0.8x(L-x)/L^2 + 0.12$             & 1       & 800, 800 & \cite{Solovjov_2017}      \\
           &                                    & $y_{CO2}=0$                                     &         &                \\
        S4 & middle third triangular to 2500    & $y_{H2O}=0.1$, $y_{CO2}=0$                      & 0.3     & 500, 500 & \cite{Solovjov_2017}      \\
        S5 & $T(x) = 1000 + 500\cos(\pi x/L)$ 	 & $y_{H2O}=0.1$, $y_{CO2}=0$                      & 2       & 1500, 500 & \cite{Solovjov_2017}     \\
        B3 & $T(x) = 400 + 1400\sin(\pi x/L)^2$ & $y_{H2O}(x) = 0.0001 + 0.9999\sin(\pi x/L)^2$	& 1       & 400, 400 & \cite{Bordbar_2020}      \\
           &                                    & $y_{CO2}=1-y_{H2O}$                             	&         &                \\
        Sb1& $T = 1000$                         & $y_{H2O} = 0.2$, $y_{CO2}=0.1$, $y_{CO}=0.03$   	& 1       & cold, cold & \cite{Solovjov_2001}    \\
        \hline
    \end{tabular}
    }
\end{table}
%

@@

Figures~\ref{f:examples} and~\ref{f:exSb1} display comparative results for the PM, WSGG, and RCSLW models implemented by RadLib in the example case configurations summarized in Table~\ref{t:examples}. In all examples, the RCSLW model data is computed using four gray gases ($n=4$) and one clear gas for consistent comparison to the WSGG model, with the exception of Example S5, in which the number of gray gases in the RCSLW model is varied. The RCSLW model is initialized using the mean temperature and composition on the domain for all cases except Example S5, which uses the maximum temperature instead. 
%
\begin{figure}
    \begin{center}
    \begin{tabular}{c c}
        Example S1 & Example S2 \\
        \includegraphics[width=2.75 in]{fig_ex_S1.pdf} &
        \includegraphics[width=2.75 in]{fig_ex_S2b.pdf} \\
        Example S3 & Example S4 \\
        \includegraphics[width=2.75 in]{fig_ex_S3a.pdf} &
        \includegraphics[width=2.75 in]{fig_ex_S4a.pdf} \\
        Example S5 & Example B3 \\
        \includegraphics[width=2.75 in]{fig_ex_S5c.pdf} &
        \includegraphics[width=2.75 in]{fig_ex_B3.pdf}
    \end{tabular}
    \caption{Results for Examples S1-S5 and B3 summarized in Table~\ref{t:examples}.}
    \label{f:examples}
    \end{center}
\end{figure}
%
\begin{figure}
    \begin{center}
        \includegraphics[width=5.5 in]{fig_ex_Sb1.pdf}
    \caption{Results for Example Sb1 summarized in Table~\ref{t:examples}.}
    \label{f:exSb1}
    \end{center}
\end{figure}
%

In general, the PM model performs poorly compared to the WSGG and RCSLW models and the LBL data. The data for Examples S1, S3, and B3 show that the PM model tends to exaggerate the trends displayed by the other models; this is also true of Example S4, though the effect is somewhat overshadowed by the WSGG model results. The PM data mimics the shape of the other curves in Example S5 as well, but with such greatly exaggerated peaks and valleys that its data was omitted from the plot to more clearly display the behavior of the WSGG and RCSLW models; for the PM model in Example S5, $Q$ varies from around -80 kW/m$^3$ at $x=0$ to a peak of 200 kW/m$^3$ at x=$1.5$ m and drops to 100 kW/m$^3$ at $x=2$ m. 
In Example S2, the PM value of $q(L)/\sigma T^4$ is off the scale at an essentially constant value of unity. The PM absorption coefficient is 27.4 atm$^{-1}$m$^{-1}$, which results in a calculated optical thicknesses of 0.09 and 0.36 m in the thick and thin layers, respectively. These values are small compared to the domain sizes considered in Example S2, which coincides with the PM model's use of the optically thin approximation, but the high error in its predictions of radiative heat flux relative to the LBL data indicates that neither the optically thin approximation nor absorption coefficients calculated with this model are accurate for this case. Example S3 is the clearest illustration of the PM model's deficiencies as discussed in Section~\ref{s:planckmean}: the PM model severely overpredicts radiative heat loss at the high temperatures that characterize most combustion processes.

The WSGG model gives somewhat accurate but inconsistent results for the example cases considered here. In Examples S1, S3, S5, B3, and Sb1, it behaves as expected, producing results that are closer to the LBL data than those produced by the PM model but fall short of the RCSLW model's level of accuracy. In general, the WSGG model appears to perform best at intermediate temperatures and deviates from the desired behavior at boundary conditions, but this is not universal. For example, the WSGG model shows uncharacteristically high accuracy in Example S2; its results with four gray gases are comparable in accuracy to the RCSLW model's performance using eight or more gray gases. In contrast, the WSGG model performs exceptionally poorly in Example S4, in which it deviates from the LBL data by nearly an order of magnitude at its largest. 
The WSGG model, alongside with the PM model, is primarily empirical and contains little to no representative physics, which could explain the unexpected behavior in Examples S2 and S4. Given that all of the example cases use parameters within those specified for the WSGG model in the literature \cite{Bordbar_2014,Bordbar_2020}, neither anomalous result has an obvious source. While we cannot offer an explanation for the WSGG model's inconsistent performance, it serves to underscore the importance of individual model validation on a case-by-case basis, a task potentially made easier with tools like RadLib.

The RCSLW model displays the most accurate results of the three implemented property models when compared to LBL data, capturing the relevant trends without exaggerating them. It is, however, the most difficult of the three models to implement and requires either correlated or tabulated spectral data to draw from. The RCSLW model can also be sensitive to the number of gray gases used, illustrated in Figure~\ref{f:examples} by Examples S2 and S5. The difference in accuracy between $n=2$ and $n=8$ gray gases is very small for all of the example cases and typically falls within the margin of error represented by the WSGG model's deviation from the LBL data. Example S2 shows the most prominent variations in predicted radiative heat flux based on changes in the number of gray gases, improving drastically in agreement to the LBL data with the addition of just four gray gases. RCSLW model convergence can also vary with the value of the chosen reference temperature; in Example S5, for instance, the RCSLW model converges to the LBL data if it is initialized with the maximum rather than the average temperature on the domain. When treated properly, however, the RCSLW model provides the most accurate results of all of the radiation property models considered here.

Figure~\ref{f:exSb1} shows the results from Example Sb1, which considers three magnitudes of the soot volume fraction in addition to the radiating gas. Both the WSGG and RCSLW models perform accurately in all three cases, but the PM model displays significant deviations from the LBL data. In sooting flames, soot acts as a significant source and sink of radiation, and its radiative properties differ significantly from those of the participating gaseous species. Soot reactions and transport are both sensitive to and strongly influence local variations in flame temperature and gas composition, forming a feedback loop in which accurate prediction of both gas and soot radiative properties becomes critical to predicting flame properties and behavior, particularly late-stage flame phenomena such as flame sheet breakthrough, smoke production, and extinction and reignition processes. The agreement produced by RadLib's approach to soot radiation in Example Sb1 is an encouraging step toward accurate, detailed treatment of soot's radiative properties in complex combustion simulations.

@@

Figure~\ref{f:fds} shows results of the simulation with comparison to experimental data. Three radiation models were included: the FDS default model, and the RCSLW and WSGG models from RadLib. 
The FDS default model uses a single gray gas with a composition and temperature dependent absorption coefficient computed from RADCAL. Further details are provided in the FDS user guide \cite{FDS}. 
Figure~\ref{f:fds}a shows radial profiles of mean temperature at two vertical positions. The models produce centerline temperature differences of about 80 K (7\%) between the RCSLW and the WSGG at z=2.5D, and 90 K (12\%) at z=3.5D. The differences are lower at lower points in the flame since the effect of radiation on temperature is cumulative. Figure~\ref{f:fds}b plots height versus radiative emission per unit length along the centerline with the spread in radiative emission around the peak values being 1.3 kW/m (11\%).
%
\begin{figure}
    \begin{center}
    \begin{tabular}{c c}
        (a) & (b) \\
        \includegraphics[width=2.5 in]{fig_fds_T_mean.pdf} &
        \includegraphics[width=2.5 in]{fig_fds_rad_mean.pdf}
    \end{tabular}
    \caption{Radial profiles of mean temperature at two heights (a), and height versus radiative emission (b) for FDS simulations of the FM Burner validation case comparing three models to experimental data.}
    \label{f:fds}
    \end{center}
\end{figure}
%

>>>>

Weighted sum of gray gases (WSGG) models generate radiation absorption coefficients for use with the spectrally integrated RTE (Eq.~\ref{e:RTE_wsgg}).
RadLib implements the WSGG model presented by Bordbar et al. \citep{Bordbar_2014,Bordbar_2020}, which uses correlations based on the HITEMP 2010 database \cite{Rothman_2010}. It models mixtures of gaseous \ce{CO2} and \ce{H2O}, the two most significant radiatively participating species in most combustion environments, using four gray gases and one clear gas. Absorption coefficients are calculated by 
%
\begin{linenomath}
\begin{equation}
    \kappa_j=\sum_{k=0}^{4}d_{j,k}M_r^k,
\end{equation}
\end{linenomath}
%
where $\kappa_j$ is the absorption coefficient for gray gas $j$, $d_{j,k}$ is a correlated model coefficient, and $M_r$ is the molar ratio $y_{\mathrm{H_2O}}/y_{\mathrm{CO_2}}$. The weight factors are calculated by 
%
\begin{linenomath}
\begin{equation}
    a_j=\sum_{k=0}^{4}b_{j,k}T_r^k,
\end{equation}
\end{linenomath}
%
where $a_j$ is the weighting factor for gas $j$ and $T_r$ is a scaled temperature equal to $T/T_{ref}$ with $T_{ref}=1200$K. The valid temperature range of the model is 300-2400 K; temperatures outside this range use the value at the respective bound. The value of $b_{i,j}$ is calculated by 
%
\begin{linenomath}
\begin{equation}
 b_{j,k}=\sum_{i=0}^{4}C_{j,k,i}M_r^i,
\end{equation}
\end{linenomath}
%
where $C_{j,k,i}$ is another correlated model coefficient. The model coefficients $d_{j,k}$ and $C_{j,k,i}$ are provided by Bordbar et~al. \citep{Bordbar_2014,Bordbar_2020} but updated values are used in RadLib.

@@

Table~\ref{t:examples} summarizes the example cases. Each case is labeled according to its source---Solvojov et al. \cite{Solovjov_2017} (S), Bordbar et al. \cite{Bordbar_2020} (B), or Solovjov et~al. \cite{Solovjov_2001} (Sb)---and the number of each example corresponds to the example number in its respective reference. Each example case consists of one or more gas layers ("slabs") bounded by black walls. 
Example S1 consists of two adjacent gas layers with identical species compositions: one hot layer of constant thickness and one cold layer of varying thickness. 
In Example S2, the two gas layers are isothermal but differ in the value of the \ce{CO2} mole fraction, where the layer thickness of the ``thin'' slab (with the smaller \ce{CO2} mole fraction) varies.
Example S3 applies parabolic temperature and \ce{H2O} mole fraction profiles across a single gas layer. 
Example S4 divides the gas into three equally-spaced layers and applies a triangular temperature profile to the middle layer and a constant temperature to the two outer layers.
Example S5 uses a half-sinusoid temperature profile that decreases from 1500 to 500 K applied to a single gas layer of constant species composition.
Example B3 has symmetric temperature and H$_2$O mole fraction profiles with central peaks of 1800 K and 1, respectively (with $y_{CO2}=1-y_{H2O}$). 
Example Sb1 consists of a single gas with a uniform temperature and composition profile and soot radiation. 
%
\begin{table}
    \caption{Summary of example cases. All cases use $P=1$ atm and black walls.}
    \label{t:examples}
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c l l c c c c}
        \hline
        Example & T(K)                          & $y_{H2O}$  $y_{CO2}$ (mole frac.)               & L (m)   & $T_{walls}$ (K) & Ref.\\
        \hline
        S1 & $T(x<0.5)=2000$; $T(x>0.5)=300$    & $y_{CO2}=0.1$, $y_{H2O}=0.2$                    & 0.5-2.5 & cold, cold	& \cite{Solovjov_2017}     \\
        S2 & T=1000                             & $y_{CO2}(x<0.5)=0.4$, $y_{CO2}(x>0.5)=0.1$      & 0.5-2.5 & cold, cold & \cite{Solovjov_2017}    \\
           &                                    & $y_{H2O}=0.0$                                   &         &                \\
        S3 & $T(x) = 4000x(L-x)/L^2 + 800$      & $y_{H2O}(x) = 0.8x(L-x)/L^2 + 0.12$             & 1       & 800, 800 & \cite{Solovjov_2017}      \\
           &                                    & $y_{CO2}=0$                                     &         &                \\
        S4 & middle third triangular to 2500    & $y_{H2O}=0.1$, $y_{CO2}=0$                      & 0.3     & 500, 500 & \cite{Solovjov_2017}      \\
        S5 & $T(x) = 1000 + 500\cos(\pi x/L)$ 	 & $y_{H2O}=0.1$, $y_{CO2}=0$                      & 2       & 1500, 500 & \cite{Solovjov_2017}     \\
        B3 & $T(x) = 400 + 1400\sin(\pi x/L)^2$ & $y_{H2O}(x) = 0.0001 + 0.9999\sin(\pi x/L)^2$	& 1       & 400, 400 & \cite{Bordbar_2020}      \\
           &                                    & $y_{CO2}=1-y_{H2O}$                             	&         &                \\
        Sb1& $T = 1000$                         & $y_{H2O} = 0.2$, $y_{CO2}=0.1$, $y_{CO}=0.03$   	& 1       & cold, cold & \cite{Solovjov_2001}    \\
        \hline
    \end{tabular}
    }
\end{table}
%

@@

Figures~\ref{f:examples} and~\ref{f:exSb1} display comparative results for the PM, WSGG, and RCSLW models implemented by RadLib in the example case configurations summarized in Table~\ref{t:examples}. These plots are included to demonstrate the examples that are provided with the code, and to compare the models to give an indication of their relative accuracy. In all examples, the RCSLW model data is computed using four gray gases ($n=4$) and one clear gas for consistent comparison to the WSGG model, with the exception of Example S5, in which the number of gray gases in the RCSLW model is varied. The RCSLW model is initialized using the mean temperature and composition on the domain for all cases except Example S5, which uses the maximum temperature. 
%
\begin{figure}
    \begin{center}
    \begin{tabular}{c c}
        Example S1 & Example S2 \\
        \includegraphics[width=2.75 in]{fig_ex_S1.pdf} &
        \includegraphics[width=2.75 in]{fig_ex_S2b.pdf} \\
        Example S3 & Example S4 \\
        \includegraphics[width=2.75 in]{fig_ex_S3a.pdf} &
        \includegraphics[width=2.75 in]{fig_ex_S4a.pdf} \\
        Example S5 & Example B3 \\
        \includegraphics[width=2.75 in]{fig_ex_S5c.pdf} &
        \includegraphics[width=2.75 in]{fig_ex_B3.pdf}
    \end{tabular}
    \caption{Results for Examples S1-S5 and B3 summarized in Table~\ref{t:examples}.}
    \label{f:examples}
    \end{center}
\end{figure}
%
\begin{figure}
    \begin{center}
        \includegraphics[width=5.5 in]{fig_ex_Sb1.pdf}
    \caption{Results for Example Sb1 summarized in Table~\ref{t:examples}.}
    \label{f:exSb1}
    \end{center}
\end{figure}
%

In general, the PM model performs poorly compared to the WSGG and RCSLW models and the LBL data. The results show that the PM model tends to exaggerate the trends displayed by the other models. The PM data mimics the shape of the other curves in Example S5 as well, but the curve is off-scale and omitted for clarity. For the PM model in Example S5, $Q$ varies from around -80 kW/m$^3$ at $x=0$ to a peak of 200 kW/m$^3$ at x=$1.5$ m and drops to 100 kW/m$^3$ at $x=2$ m. 
In Example S2, the PM value of $q(L)/\sigma T^4$ is off the scale at an essentially constant value of unity. The PM absorption coefficient is 27.4 atm$^{-1}$m$^{-1}$, which results in a calculated optical thicknesses of 0.09 and 0.36 m in the thick and thin layers, respectively. These values are small compared to the domain sizes greater than 0.5 m that are considered in Example S2.
%, which coincides with the PM model's use of the optically thin approximation, but the high error in its predictions of radiative heat flux relative to the LBL data indicates that neither the optically thin approximation nor absorption coefficients calculated with this model are accurate for this case. 
Example S3 is the clearest illustration of the PM model's deficiencies as discussed in Section~\ref{s:planckmean}: the PM model overpredicts radiative heat loss at the high temperatures that characterize most combustion processes.

The WSGG model closely follows the trends of the LBL data for all examples, and in most examples gives reasonable quantitative agreement. This is especially true for Example S2, S4, and B3. For example S5, the WSGG model shows peak errors of around 50\%, and somewhat higher in Example S1. The heat flux near the boundaries in Example B3 is low by around 30\%. The RCSLW model gives results in very close agreement to the LBL data for all examples. A notable exception is example S2, where the heat flux at the right boundary is not well-predicted for four gray gases. The heat flux does converge to the LBL results when the number of gray gases increases. This behavior was not discussed in \cite{Solovjov_2017} where 24 gray gases were used. The WSGG model is very close to the LBL data for this case. The good agreement is likely a result of the simplicity of this case that is isothermal with pure \ce{CO2} at uniform concentration in either of the two adjacent gas slabs.
The RCSLW model is sensitive to the number of gray gases used, as illustrated in Figure~\ref{f:examples} by Examples S2 and S5. The difference in accuracy between $n=2$ and $n=8$ gray gases is small for all of the example cases and typically falls within the margin of error represented by the WSGG model's deviation from the LBL data. 
The RCSLW model also depends on the value of the chosen reference temperature that is used to set the ALBDF grid. In Example S5, the RCSLW model does not converge to the LBL data at $x>0.5$ m when the domain average temperature was used.

%The WSGG model gives somewhat accurate but inconsistent results for the example cases considered here. In Examples S1, S3, S5, B3, and Sb1, it behaves as expected, producing results that are closer to the LBL data than those produced by the PM model but fall short of the RCSLW model's level of accuracy. In general, the WSGG model appears to perform best at intermediate temperatures and deviates from the desired behavior at boundary conditions, but this is not universal. For example, the WSGG model shows uncharacteristically high accuracy in Example S2; its results with four gray gases are comparable in accuracy to the RCSLW model's performance using eight or more gray gases. In contrast, the WSGG model performs exceptionally poorly in Example S4, in which it deviates from the LBL data by nearly an order of magnitude at its largest. 
%The WSGG model, alongside with the PM model, is primarily empirical and contains little to no representative physics, which could explain the unexpected behavior in Examples S2 and S4. Given that all of the example cases use parameters within those specified for the WSGG model in the literature \cite{Bordbar_2014,Bordbar_2020}, neither anomalous result has an obvious source. While we cannot offer an explanation for the WSGG model's inconsistent performance, it serves to underscore the importance of individual model validation on a case-by-case basis, a task potentially made easier with tools like RadLib.

%The RCSLW model displays the most accurate results of the three implemented property models when compared to LBL data, capturing the relevant trends without exaggerating them. It is, however, the most difficult of the three models to implement and requires either correlated or tabulated spectral data to draw from. 
%The RCSLW model can be sensitive to the number of gray gases used, illustrated in Figure~\ref{f:examples} by Examples S2 and S5. The difference in accuracy between $n=2$ and $n=8$ gray gases is small for all of the example cases and typically falls within the margin of error represented by the WSGG model's deviation from the LBL data. Example S2 shows the most prominent variations in predicted radiative heat flux based on changes in the number of gray gases, improving drastically in agreement to the LBL data with the addition of just four gray gases. RCSLW model convergence can also vary with the value of the chosen reference temperature; in Example S5, for instance, the RCSLW model converges to the LBL data if it is initialized with the maximum rather than the average temperature on the domain. When treated properly, however, the RCSLW model provides the most accurate results of all of the radiation property models considered here.

Figure~\ref{f:exSb1} shows the results from Example Sb1, which considers three magnitudes of the soot volume fraction in addition to the radiating gas. Both the WSGG and RCSLW models perform accurately in all three cases, but the PM model displays significant deviations from the LBL data. In sooting flames, soot acts as a significant source and sink of radiation, and its radiative properties differ significantly from those of the participating gaseous species. Soot reactions and transport are both sensitive to and strongly influence local variations in flame temperature and gas composition, forming a feedback loop in which accurate prediction of both gas and soot radiative properties becomes critical to predicting flame properties and behavior, particularly late-stage flame phenomena such as flame sheet breakthrough, smoke production, and extinction and reignition processes.
%The agreement produced by RadLib's approach to soot radiation in Example Sb1 is an encouraging step toward accurate, detailed treatment of soot's radiative properties in complex combustion simulations.

@@

Figure~\ref{f:fds} shows results of the simulation with comparison to experimental data. Three radiation models were included: the FDS default model, and the RCSLW and WSGG models from RadLib (both using four gray gases and one clear gas). 
The FDS default model uses a single gray gas with a composition and temperature dependent absorption coefficient computed from RADCAL. Further details are provided in the FDS user guide \cite{FDS}. 
Figure~\ref{f:fds}a shows radial profiles of mean temperature at two vertical positions. The models produce centerline temperature differences of about 80 K (7\%) between the RCSLW and the WSGG at z=2.5D, and 90 K (12\%) at z=3.5D. The differences are lower at lower points in the flame since the effect of radiation on temperature is cumulative. Figure~\ref{f:fds}b plots height versus radiative emission per unit length along the centerline with the spread in radiative emission around the peak values being 1.3 kW/m (11\%).
%
\begin{figure}
    \begin{center}
    \begin{tabular}{c c}
        (a) & (b) \\
        \includegraphics[width=2.5 in]{fig_fds_T_mean.pdf} &
        \includegraphics[width=2.5 in]{fig_fds_rad_mean.pdf}
    \end{tabular}
    \caption{Radial profiles of mean temperature at two heights (a), and height versus radiative emission (b) for FDS simulations of the FM Burner validation case comparing three models to experimental data.}
    \label{f:fds}
    \end{center}
\end{figure}
%
